---
title: "Dataset cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Step 1. Get data from the World Bank API

```{r}
install.packages("WDI")
```

```{r}
library('WDI')
```

```{r}
WDIsearch('gdp')
```

```{r}
WDIsearch('external.*debt')
```

```{r}
WDIsearch('exchange.*rate')
```

```{r}
Data_WDI = WDI(indicator=c('NEER', 'NY.GDP.PCAP.KD','FI.RES.TOTL.DT.ZS','DT.DOD.DECT.GN.ZS'), country='all', start=1990, end=2018)
head(Data_WDI)
```

## Name of Variables in current dataframe
'NEER'                 Nominal effective exchange rate
'NY.GDP.PCAP.KD'       GDP per capita (constant 2010 US$)                                                                                     
'FI.RES.TOTL.DT.ZS',   Total reserves (% of total external debt)
'DT.DOD.DECT.GN.ZS',   External debt stocks (% of GNI)

```{r}
names(Data_WDI) <- c('iso2c', 'country', 'year', 'nominal.exchange.rate', 'GDP.per.capita', 'total.reserves.outof.total.external.debt', 'external.Debt/GNI')
```

##GET DATA ON LEVEL OF DEMOCRACY ----- data is downloaded from Center for Systemic Peace website.

```{r}
getwd()
```

```{r}
Data_polity <- read.csv('/Users/liuminchen/Documents/Currency Crisis/DATASET/politydata.csv')
```

##MERGE WDI AND POLITYDATA SETS

```{r}
WDI_POLITY <- merge(Data_WDI,Data_polity, by=c("iso2c","year"))
```

```{r}
head(WDI_POLITY)
```

##GET DATA FROM THE IMF ON US 3-MONTH TREASURY BILL YIELD

```{r}
install.packages("IMFData")
```

```{r}
library(IMFData)
```


```{r}
IFS.available.codes <- DataStructureMethod('IFS') # Get dimension code of IFS dataset
names(IFS.available.codes) # Available dimension code
IFS.available.codes[[1]] # Possible code in the first dimension
CodeSearch(IFS.available.codes, "CL_INDICATOR_IFS", "Treasury")
```

```{r}
queryfilter <- list(CL_FREA = "A", CL_AREA_IFS = "US", CL_INDICATOR_IFS = "FITB_3M_PA")
```


```{r}
databaseID <- "IFS"
startdate = "1980-01-01"
enddate = "2018-12-31"
checkquery = FALSE
```

```{r}
USTREASURY <- CompactDataMethod(databaseID, queryfilter, startdate, enddate, 
    checkquery, tidy = TRUE)
head(USTREASURY)
```


## Merge the dataset

```{r}
library(tidyverse)
```


```{r}
USTREASURY2 <- USTREASURY[,c(1,2)]
head(USTREASURY2)
```

```{r}
names(USTREASURY2) <- c('year', 'US.3m.treasury.yield')
```

```{r}
WDI_POLITY_USTREASURY <- merge(WDI_POLITY, USTREASURY2, by="year")
head(WDI_POLITY_USTREASURY)
```

```{r}
WDI_POLITY_USTREASURY <- WDI_POLITY_USTREASURY[order( WDI_POLITY_USTREASURY[,2] ),]
WDI_POLITY_USTREASURY2 <- WDI_POLITY_USTREASURY[,-c(8,9)]
head(WDI_POLITY_USTREASURY2)
```

## GENERATE DEPENDENT VARIABLE (DUMMY)
```{r}
WDI_POLITY_USTREASURY2$currency.crisis = ifelse((WDI_POLITY_USTREASURY2$nominal.exchange.rate - lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate))/lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate) <= -0.15, 1, 0)
```

## GENERATE INDEPENDENT LAG VARIABLES
```{r}
WDI_POLITY_USTREASURY2$lag.GDP.per.capita = lag(WDI_POLITY_USTREASURY2$GDP.per.capita)
WDI_POLITY_USTREASURY2$lag.total.reserves = lag(WDI_POLITY_USTREASURY2$total.reserves.outof.total.external.debt)
```

```{r}
View(WDI_POLITY_USTREASURY2)
```

```{r}
mean(WDI_POLITY_USTREASURY2$currency.crisis)
```

```{r}
data_clean <- WDI_POLITY_USTREASURY2 %>% drop_na(currency.crisis)
View(data_clean)
```

```{r}
mean(data_clean$currency.crisis)
```

## CLEAN NULL VALUES IN THE COMPLETE MERGED DATASET
```{r}
isNA <- apply(is.na(data_clean), 1, any)
data_clean <- data_clean[!isNA, ]
```

```{r}
mean(data_clean$currency.crisis)
```

## SAVE THE COMPLETE MERGED DATASET.
```{r}
write.csv(data_clean,'data_clean_for_next_steps.csv')
```

