---
title: "Final Work"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Step 1. Get data from the World Bank API

```{r}
install.packages("WDI")
```

```{r}
library('WDI')
```

```{r}
WDIsearch('gdp')
```

```{r}
WDIsearch('external.*debt')
```

```{r}
WDIsearch('exchange.*rate')
```

```{r}
Data_WDI = WDI(indicator=c('NEER', 'NY.GDP.PCAP.KD','FI.RES.TOTL.DT.ZS','DT.DOD.DECT.GN.ZS'), country='all', start=1990, end=2018)
head(Data_WDI)
```

## Name of Variables in current dataframe
'NEER'                 Nominal effective exchange rate
'NY.GDP.PCAP.KD'       GDP per capita (constant 2010 US$)                                                                                     
'FI.RES.TOTL.DT.ZS',   Total reserves (% of total external debt)
'DT.DOD.DECT.GN.ZS',   External debt stocks (% of GNI)

```{r}
names(Data_WDI) <- c('iso2c', 'country', 'year', 'nominal.exchange.rate', 'GDP.per.capita', 'total.reserves.outof.total.external.debt', 'external.Debt/GNI')
```

##GET DATA ON LEVEL OF DEMOCRACY ----- data is downloaded from Center for Systemic Peace website.

```{r}
getwd()
```

```{r}
Data_polity <- read.csv('/Users/liuminchen/Documents/Currency Crisis/DATASET/politydata.csv')
```

##MERGE WDI AND POLITYDATA SETS

```{r}
WDI_POLITY <- merge(Data_WDI,Data_polity, by=c("iso2c","year"))
```

```{r}
head(WDI_POLITY)
```

##GET DATA FROM THE IMF ON US 3-MONTH TREASURY BILL YIELD

```{r}
install.packages("IMFData")
```

```{r}
library(IMFData)
```


```{r}
IFS.available.codes <- DataStructureMethod('IFS') # Get dimension code of IFS dataset
names(IFS.available.codes) # Available dimension code
IFS.available.codes[[1]] # Possible code in the first dimension
CodeSearch(IFS.available.codes, "CL_INDICATOR_IFS", "Treasury")
```

```{r}
queryfilter <- list(CL_FREA = "A", CL_AREA_IFS = "US", CL_INDICATOR_IFS = "FITB_3M_PA")
```


```{r}
databaseID <- "IFS"
startdate = "1980-01-01"
enddate = "2018-12-31"
checkquery = FALSE
```

```{r}
USTREASURY <- CompactDataMethod(databaseID, queryfilter, startdate, enddate, 
    checkquery, tidy = TRUE)
head(USTREASURY)
```


## Merge the dataset

```{r}
library(tidyverse)
```


```{r}
USTREASURY2 <- USTREASURY[,c(1,2)]
head(USTREASURY2)
```

```{r}
names(USTREASURY2) <- c('year', 'US.3m.treasury.yield')
```

```{r}
WDI_POLITY_USTREASURY <- merge(WDI_POLITY, USTREASURY2, by="year")
head(WDI_POLITY_USTREASURY)
```

```{r}
WDI_POLITY_USTREASURY <- WDI_POLITY_USTREASURY[order( WDI_POLITY_USTREASURY[,2] ),]
WDI_POLITY_USTREASURY2 <- WDI_POLITY_USTREASURY[,-c(8,9)]
head(WDI_POLITY_USTREASURY2)
```

## GENERATE DEPENDENT VARIABLE (DUMMY)
```{r}
WDI_POLITY_USTREASURY2$currency.crisis = ifelse((WDI_POLITY_USTREASURY2$nominal.exchange.rate - lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate))/lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate) <= -0.15, 1, 0)
```

## GENERATE INDEPENDENT LAG VARIABLES
```{r}
WDI_POLITY_USTREASURY2$lag.GDP.per.capita = lag(WDI_POLITY_USTREASURY2$GDP.per.capita)
WDI_POLITY_USTREASURY2$lag.total.reserves = lag(WDI_POLITY_USTREASURY2$total.reserves.outof.total.external.debt)
```

```{r}
View(WDI_POLITY_USTREASURY2)
```

```{r}
mean(WDI_POLITY_USTREASURY2$currency.crisis)
```

```{r}
data_clean <- WDI_POLITY_USTREASURY2 %>% drop_na(currency.crisis)
View(data_clean)
```

```{r}
mean(data_clean$currency.crisis)
```

## CLEAN NULL VALUES IN THE COMPLETE MERGED DATASET
```{r}
isNA <- apply(is.na(data_clean), 1, any)
data_clean <- data_clean[!isNA, ]
```

```{r}
mean(data_clean$currency.crisis)
```


```{r}
names(data_clean) <- c('year', 'isoc2', 'countryx', 'nominalexchangerate', 'gdppercapita', 'totalreservesoutoftotalexternald', 'externaldebtgni', 'polity2', 'us3mtreasuryyield', 'currencycrisis', 'laggdppercapita', 'lagtotalreserves')
```

```{r}
View(data_clean)
```


## SAVE THE COMPLETE MERGED DATASET.
```{r}
write.csv(data_clean,'data_clean_for_next_steps.csv')
```

###Data visualization 
Display the connections between Dependent variable (currency crisis) and control variables, or the connection among control variables.

```{r}
require(skimr)
```

```{r}
require(readr)
require(tidyverse)
require(ggplot2)
require(dplyr)
```


```{r}
df <- read.csv('C:/Users/Meng/Desktop/Spring2019/DataScience/group project/new project/New-Group-Project/data_clean_for_next_steps.csv')
```

```{r}
head(df)

```


summary of statistics:
describe the variables
```{r}
skimr::skim(df)
```


#creat df1 to select only China, Argentina, Brazil, Romania, Russia Federation, Sierra Leone, Thailand, Turkey and South Africa
df1 is used to create graph1
```{r}
df1 <- df%>%
  filter(countryx=="China"|countryx=="Brazil"|countryx=="Romania"|countryx=="Russian Federation"|countryx=="Sierra Leone"|countryx=="Thailand"|countryx=="Turkey"|countryx=="South Africa"|countryx=="Nigeria"|countryx=="Zambia"|countryx=='Ukraine')
```

```{r}
df1$currencycrisis <- factor(df1$currencycrisis)
```

display the external debt/gni by 11 countries 
```{r}
ggplot(df1,aes(year,externaldebtgni))+
  geom_line()+
  ylab("external debt/gni")+
  facet_wrap(~countryx)
```
#graph 1
Is it the higher the external debt, the more often currency crisis happened?
-yes, for Nigeria,Zambia,Thailand,Sierra Leoone.
```{r,fig.height=8,fig.width=15}
p <- ggplot(df1,aes(year,externaldebtgni))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("external debt/gni")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)
  
p
                
```
#graph 2
multicollinearity:
see the relationship between control variable: polity score and gdp per capita
display polity scores by countries:
```{r}
q1 <- ggplot(df1,aes(year,polity2))+
  geom_line()+
  ylab("polity score")+
  facet_wrap(~countryx)
q1
```
display GDP per capita by countries
```{r}
q2 <- ggplot(df1,aes(year,laggdppercapita,color="red"))+
  geom_line()+
  facet_wrap(~countryx)
q2
```

multicolinear?
No. multicolinearity doesn't seem to exist.
it make sense: Nigeria, Romania -> The more "democratic", the higher lagged gdp per capita
not make sense: China, Brazil,Russia, Ukraine, Sierra Leone, South Africa, Thailand, Turkey
1) The polity score might not a strong factor reflecting political stability
2) The polity score might not be strongly asscoiated with some countries' currency crisis.
```{r,fig.height=8,fig.width=15}
ggplot(df1,aes(x=year))+
  geom_line(aes(year,laggdppercapita),color="steel blue")+
  geom_line(aes(year,(polity2+10)*500),color="dark red")+
  scale_y_continuous(name="lagged gdp per capita", sec.axis=sec_axis(~./500-10, name="polity score"))+
  theme(axis.title.y.left=element_text(color="steel blue"),
    axis.text.y.left=element_text(color="steel blue"),
    axis.title.y.right=element_text(color="dark red"),
    axis.text.y.right=element_text(color="dark red"))+
  facet_wrap(~countryx)
```
#graph 3
external shock: How U.S. interest rate influence EMEs' currency crisis?

```{r}
 df2 <- df%>%
  filter(currencycrisis==1)
```
Count the number of currency crisis of all selected countries by years
```{r}
ggplot(df2,aes(x=year))+
geom_bar(mapping = aes(x = year, y = currencycrisis), stat = "identity") 
```
Display the U.S. interest rate by years
```{r}
ggplot(df2,aes(x=year))+
  geom_line(aes(year,us3mtreasuryyield))+
  ylab("US Interest Rate")
```
The trends of US interest rate and # of countries undergoing currency crisis appear to move in the same direction approximately. 
US interest rate as an external shock has an impact on EMEs' currency crisis.
#relationship found
```{r}
ggplot(df2,aes(x=year))+
  geom_line(aes(year,us3mtreasuryyield*2.5),color="dark red")+
  geom_bar(mapping = aes(x = year, y = currencycrisis), stat = "identity",fill="steel blue",alpha=0.4)+
  scale_y_continuous(name="# of countries with currency crisis", sec.axis=sec_axis(~./2.5, name="US interest rate"))+
  theme(axis.title.y.left=element_text(color="steel blue"),
    axis.text.y.left=element_text(color="steel blue"),
    axis.title.y.right=element_text(color="dark red"),
    axis.text.y.right=element_text(color="dark red"))
```
#graph 4
polity score and currency crisis by country: no direct relationship found
```{r,fig.height=8,fig.width=15}
 ggplot(df1,aes(year,polity2))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("polity score")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)

```
#graph 5
The lower the foreign reserves, the more currency crisis happened.
strong relationship!
```{r,fig.height=8,fig.width=15}
 ggplot(df1,aes(year,totalreservesoutoftotalexternald))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("total foreign reserves out of total external debt")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)
```

