---
title: "Final Work"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


PART I. GET THE DATASET
##Step 1. Get data from the World Bank API

```{r}
install.packages("WDI")
```

```{r}
library('WDI')
```

```{r}
WDIsearch('gdp')
```

```{r}
WDIsearch('external.*debt')
```

```{r}
WDIsearch('exchange.*rate')
```

```{r}
Data_WDI = WDI(indicator=c('NEER', 'NY.GDP.PCAP.KD','FI.RES.TOTL.DT.ZS','DT.DOD.DECT.GN.ZS'), country='all', start=1990, end=2018)
head(Data_WDI)
```

## Name of Variables in current dataframe
'NEER'                 Nominal effective exchange rate
'NY.GDP.PCAP.KD'       GDP per capita (constant 2010 US$)                                                                                     
'FI.RES.TOTL.DT.ZS',   Total reserves (% of total external debt)
'DT.DOD.DECT.GN.ZS',   External debt stocks (% of GNI)

```{r}
names(Data_WDI) <- c('iso2c', 'country', 'year', 'nominal.exchange.rate', 'GDP.per.capita', 'total.reserves.outof.total.external.debt', 'external.Debt/GNI')
```

##GET DATA ON LEVEL OF DEMOCRACY ----- data is downloaded from Center for Systemic Peace website.

```{r}
getwd()
```

```{r}
Data_polity <- read.csv('C:/Users/Meng/Desktop/Spring2019/DataScience/group project/new project/New-Group-Project/DATASET/politydata.csv')
```

##MERGE WDI AND POLITYDATA SETS

```{r}
WDI_POLITY <- merge(Data_WDI,Data_polity, by=c("iso2c","year"))
```

```{r}
head(WDI_POLITY)
```

##GET DATA FROM THE IMF ON US 3-MONTH TREASURY BILL YIELD

```{r}
install.packages("IMFData")
```

```{r}
library(IMFData)
```


```{r}
IFS.available.codes <- DataStructureMethod('IFS') # Get dimension code of IFS dataset
names(IFS.available.codes) # Available dimension code
IFS.available.codes[[1]] # Possible code in the first dimension
CodeSearch(IFS.available.codes, "CL_INDICATOR_IFS", "Treasury")
```

```{r}
queryfilter <- list(CL_FREA = "A", CL_AREA_IFS = "US", CL_INDICATOR_IFS = "FITB_3M_PA")
```


```{r}
databaseID <- "IFS"
startdate = "1980-01-01"
enddate = "2018-12-31"
checkquery = FALSE
```

```{r}
USTREASURY <- CompactDataMethod(databaseID, queryfilter, startdate, enddate, 
    checkquery, tidy = TRUE)
head(USTREASURY)
```


## Merge the dataset

```{r}
library(tidyverse)
```


```{r}
USTREASURY2 <- USTREASURY[,c(1,2)]
head(USTREASURY2)
```

```{r}
names(USTREASURY2) <- c('year', 'US.3m.treasury.yield')
```

```{r}
WDI_POLITY_USTREASURY <- merge(WDI_POLITY, USTREASURY2, by="year")
head(WDI_POLITY_USTREASURY)
```

```{r}
WDI_POLITY_USTREASURY <- WDI_POLITY_USTREASURY[order( WDI_POLITY_USTREASURY[,2] ),]
WDI_POLITY_USTREASURY2 <- WDI_POLITY_USTREASURY[,-c(8,9)]
head(WDI_POLITY_USTREASURY2)
```

## GENERATE DEPENDENT VARIABLE (DUMMY)
```{r}
WDI_POLITY_USTREASURY2$currency.crisis = ifelse((WDI_POLITY_USTREASURY2$nominal.exchange.rate - lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate))/lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate) <= -0.15, 1, 0)
```

## GENERATE INDEPENDENT LAG VARIABLES
```{r}
WDI_POLITY_USTREASURY2$lag.GDP.per.capita = lag(WDI_POLITY_USTREASURY2$GDP.per.capita)
WDI_POLITY_USTREASURY2$lag.total.reserves = lag(WDI_POLITY_USTREASURY2$total.reserves.outof.total.external.debt)
```


```{r}
mean(WDI_POLITY_USTREASURY2$currency.crisis)
```

```{r}
data_clean <- WDI_POLITY_USTREASURY2 %>% drop_na(currency.crisis)
```

```{r}
mean(data_clean$currency.crisis)
```

## CLEAN NULL VALUES IN THE COMPLETE MERGED DATASET
```{r}
isNA <- apply(is.na(data_clean), 1, any)
data_clean <- data_clean[!isNA, ]
```

```{r}
mean(data_clean$currency.crisis)
```


```{r}
names(data_clean) <- c('year', 'isoc2', 'countryx', 'nominalexchangerate', 'gdppercapita', 'totalreservesoutoftotalexternald', 'externaldebtgni', 'polity2', 'us3mtreasuryyield', 'currencycrisis', 'laggdppercapita', 'lagtotalreserves')
```

```{r}
View(data_clean)
```


## SAVE THE COMPLETE MERGED DATASET.
```{r}
write.csv(data_clean,'data_clean_for_next_steps.csv')
```

PART II. DATA VISUALIZATION
###Data visualization 
Display the connections between Dependent variable (currency crisis) and control variables, or the connection among control variables.

```{r}
require(skimr)
```

```{r}
require(readr)
require(tidyverse)
require(ggplot2)
require(dplyr)
```


```{r}
df <- read.csv('C:/Users/Meng/Desktop/Spring2019/DataScience/group project/new project/New-Group-Project/data_clean_for_next_steps.csv')
```


summary of statistics:
describe the variables
```{r}
skimr::skim(df)
```


#creat df1 to select only China, Argentina, Brazil, Romania, Russia Federation, Sierra Leone, Thailand, Turkey and South Africa
df1 is used to create graph1
```{r}
df1 <- df%>%
  filter(countryx=="China"|countryx=="Brazil"|countryx=="Romania"|countryx=="Russian Federation"|countryx=="Sierra Leone"|countryx=="Thailand"|countryx=="Turkey"|countryx=="South Africa"|countryx=='Ukraine')
```

```{r}
df1$currencycrisis <- factor(df1$currencycrisis)
```

display the external debt/gni by 11 countries 
```{r}
ggplot(df1,aes(year,externaldebtgni))+
  geom_line()+
  ylab("external debt/gni")+
  facet_wrap(~countryx)
```
#graph 1
Is it the higher the external debt, the more often currency crisis happened?
-yes, for Nigeria,Zambia,Thailand,Sierra Leoone.
```{r,fig.height=5,fig.width=10}
graph1 <- ggplot(df1,aes(year,externaldebtgni))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("external debt/gni")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)+
  theme(legend.position = "bottom")+
  labs(x="year",
       y="external debt/GNI",
       fill="",
       title="Higher external debt, More currency crisis?",
       subtitle="Russia, Sierra Leone, Thailand: Yes") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust=.5),
        plot.subtitle = element_text(hjust=.5))+
  ggsave("graph1.jpg")
graph1
```
#graph 2
multicollinearity:
see the relationship between control variable: polity score and gdp per capita
display polity scores by countries:
```{r}
q1 <- ggplot(df1,aes(year,polity2))+
  geom_line()+
  ylab("polity score")+
  facet_wrap(~countryx)
q1
```
display GDP per capita by countries
```{r}
q2 <- ggplot(df1,aes(year,laggdppercapita,color="red"))+
  geom_line()+
  facet_wrap(~countryx)
q2
```

multicolinear?
No. multicolinearity doesn't seem to exist.
it make sense: Nigeria, Romania -> The more "democratic", the higher lagged gdp per capita
not make sense: China, Brazil,Russia, Ukraine, Sierra Leone, South Africa, Thailand, Turkey
1) The polity score might not a strong factor reflecting political stability
2) The polity score might not be strongly asscoiated with some countries' currency crisis.
```{r,fig.height=8,fig.width=15}
graph2 <- ggplot(df1,aes(x=year))+
  geom_line(aes(year,laggdppercapita),color="steel blue")+
  geom_line(aes(year,(polity2+10)*500),color="dark red")+
  scale_y_continuous(name="lagged gdp per capita", sec.axis=sec_axis(~./500-10, name="polity score"))+
  theme(axis.title.y.left=element_text(color="steel blue"),
    axis.text.y.left=element_text(color="steel blue"),
    axis.title.y.right=element_text(color="dark red"),
    axis.text.y.right=element_text(color="dark red"))+
  facet_wrap(~countryx)+
  theme(legend.position = "bottom")+
  labs(title="Higher Polity Score, Higher GDP per capita?",
       subtitle="Not necessary") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust=.5),
        plot.subtitle = element_text(hjust=.5))+
  ggsave("graph2.jpg")
graph2
```
#graph 3
external shock: How U.S. interest rate influence EMEs' currency crisis?

```{r}
 df2 <- df%>%
  filter(currencycrisis==1)
```
Count the number of currency crisis of all selected countries by years
```{r}
ggplot(df2,aes(x=year))+
geom_bar(mapping = aes(x = year, y = currencycrisis), stat = "identity") 
```
Display the U.S. interest rate by years
```{r}
ggplot(df2,aes(x=year))+
  geom_line(aes(year,us3mtreasuryyield))+
  ylab("US Interest Rate")
```
The trends of US interest rate and # of countries undergoing currency crisis appear to move in the same direction approximately. 
US interest rate as an external shock has an impact on EMEs' currency crisis.
#relationship found
```{r}
graph3 <- ggplot(df2,aes(x=year))+
  geom_line(aes(year,us3mtreasuryyield*2.5),color="dark red")+
  geom_bar(mapping = aes(x = year, y = currencycrisis), stat = "identity",fill="steel blue",alpha=0.4)+
  scale_y_continuous(name="# of countries with currency crisis", sec.axis=sec_axis(~./2.5, name="US interest rate"))+
  theme(axis.title.y.left=element_text(color="steel blue"),
    axis.text.y.left=element_text(color="steel blue"),
    axis.title.y.right=element_text(color="dark red"),
    axis.text.y.right=element_text(color="dark red"))+
  labs(title="How U.S. interest rate affect currency crisis in EMEs") +
  theme(plot.title = element_text(hjust=.5),
        plot.subtitle = element_text(hjust=.5))+
  ggsave("graph3.jpg")
graph3
```
#graph 4
polity score and currency crisis by country: no direct relationship found
```{r,fig.height=8,fig.width=15}
graph4 <- ggplot(df1,aes(year,polity2))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("polity score")+
  scale_color_manual(values=c("black","red"))+
  labs(title="Higher polity score, More currency crisis?") +
  theme(legend.position = "bottom",
    plot.title = element_text(hjust=.5),
        plot.subtitle = element_text(hjust=.5))+
  facet_wrap(~countryx)+
  ggsave("graph4.jpg")
graph4

```
#graph 5
The lower the foreign reserves, the more currency crisis happened.
strong relationship!
```{r,fig.height=8,fig.width=15}
graph5 <- ggplot(df1,aes(year,totalreservesoutoftotalexternald))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("total foreign reserves out of total external debt")+
  scale_color_manual(values=c("black","red"))+
  labs(title="Lower foreign reserves, more currenct crisis!") +
  theme(legend.position = 'bottom',
    plot.title = element_text(hjust=.5),
        plot.subtitle = element_text(hjust=.5))+
  facet_wrap(~countryx)+
  ggsave("graph5.jpg")
graph5
```
merge graph1, graph3, graph5 to see how debt, U.S. interest rate, and foreign reserves associated with currency crisis in EMEs
```{r}
install.packages('ggpubr')
```

```{r}
library(ggpubr)
```

```{r,fig.height=10,fig.width=15}
a <- ggarrange(graph1,graph5,ncol = 2, nrow = 1)+
  ggsave("merge15.jpg")
a
```


```{r,fig.height=10,fig.width=8}
b <- ggarrange(graph2,graph4,ncol = 1, nrow = 2)+
  ggsave("merge24.jpg")
b
```

PART III. LOGIT REGRESSION

```{r}
final_data <- read.csv('data_clean_for_next_steps.csv')
```

```{r}
install.packages("margins")
install.packages("ggeffects")
```



```{r}
library(ggeffects)
```

```{r}
library("lme4")
```


```{r}
m <- glmer(currencycrisis ~ laggdppercapita + lagtotalreserves + externaldebtgni + polity2 + us3mtreasuryyield + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m, corr = FALSE)
```
##To Tingjie: use a table to visualize it
```{r}
summary(m)


```

```{r}
library("texreg")
```

```{r}
table1 <- screenreg(m)
table1
```




```{r}
plot(fitted(m), residuals(m), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
```



```{r}
margin_eff <- ggpredict(m, term = "externaldebtgni")
margin_eff
```

##To Tingjie: margin_eff 1,2,3,4,5 need to be printed
```{r}
partial1 <- plot(margin_eff)
partial1
```

```{r}
margin_eff2 <- ggpredict(m, term = "lagtotalreserves")
margin_eff2
```

```{r}
partial2 <- plot(margin_eff2)
partial2
```

```{r}
margin_eff3 <- ggpredict(m, term = "us3mtreasuryyield")
margin_eff3
```

```{r}
partial3 <- plot(margin_eff3)
partial3
```

```{r}
margin_eff4 <- ggpredict(m, term = "laggdppercapita")
margin_eff4
```

```{r}
partial4 <- plot(margin_eff4)
partial4
```


```{r}
margin_eff5 <- ggpredict(m, term = "polity2")
margin_eff5
```

```{r}
partial5 <- plot(margin_eff5)
partial5
```
#Merged 
```{r,fig.width=14,fig.height=8}
partial <- ggarrange(partial1,partial2,partial3,partial4,partial5)

partialmerge <- annotate_figure(partial,
                top = text_grob("Partial dependency plot", color = "dark grey", face = "bold", size = 18))+
  ggsave("partialmerge.jpg")
partialmerge
```



```{r}
m3 <- glmer(currencycrisis ~  externaldebtgni + polity2 + us3mtreasuryyield + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m3, corr = FALSE)
summary(m3)
```

```{r}
m4 <- glmer(currencycrisis ~  lagtotalreserves + externaldebtgni + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m4, corr = FALSE)
summary(m4)
```

```{r}
m5 <- glmer(currencycrisis ~  externaldebtgni + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m5, corr = FALSE)
summary(m5)
```


### PART III. MACHINE LEARNING 

## read the dataset
```{r}
getwd()
```

```{r}
ml_data <- read.csv(file = "C:/Users/Meng/Desktop/Spring2019/DataScience/group project/new project/New-Group-Project/data_clean_for_next_steps.csv")
```


## Task 1. We split the data into a training and test dataset, partitioning data before 2014 into the training data, and holding out data in and after 2014 as a test set.
```{r}
install.packages("caret")
```

```{r}
library(caret)
library(dplyr)
library(tidyverse)
```

```{r}
test = ml_data %>% filter(year >= 2014)
train = ml_data %>% filter(year < 2014)
```

```{r}
View(test)
```

## step_knnimpute(all_predictors())%>% --- impute missing value

```{r}
library(recipes)
```

## convert factor into numeric
```{r}
train2 <- 
  train %>% mutate_if(is.factor, as.numeric)
head(train2)
```

```{r}
test2 <- 
  test %>% mutate_if(is.factor, as.numeric)
```

## normalize the scale
```{r}
rcp <-
  recipe(countryx~.,train2) %>%
  step_range(countryx, polity2,us3mtreasuryyield,laggdppercapita, lagtotalreserves, externaldebtgni) %>%  # Normalize scale
  prep()
# Apply the recipe to the training and test data
train3 <- bake(rcp,train2)
test3 <- bake(rcp,test2)
View(test3)
View(train3)
```

## keep variables we need for the ML model
```{r}
train4 <- train3[,c(4,8,9,10,11,12,13)]
test4 <- test3[,c(4,8,9,10,11,12,13)]
View(test4)
View(train4)
```

## Task 4. Cross-Validation

```{r}
set.seed(1988) # set a seed for replication purposes 
folds <- createFolds(train4$currencycrisis, k = 5) # Partition the data into 10 equal folds
sapply(folds,length)
```

```{r}
control_conditions <- 
  trainControl(method='cv', # K-fold cross validation
               summaryFunction = twoClassSummary, # Need this b/c it's a classification problem
               classProbs = TRUE, # Need this b/c it's a classification problem
               index = folds # The indices for our folds (so they are always the same)
  )
```




## Task 5. Models
# 1. KNN

```{r}
train4$currencycrisis <- factor(train4$currencycrisis)
test4$currencycrisis <- factor(test4$currencycrisis)

```

```{r}
levels(train4$currencycrisis) <- c("no", "yes")
levels(test4$currencycrisis) <- c("no", "yes")
```

```{r}
mod_knn <-
  train(currencycrisis ~ ., # Equation (outcome and everything else)
        data=train4, # Training data 
        method = "knn", # K-Nearest Neighbors Algorithm
        metric = "ROC", # area under the curve
        trControl = control_conditions
  )
mod_knn
```
```{r}
knn_tune = expand.grid(k = c(1,3,5,6,7,10))
knn_tune
```

```{r}
mod_knn2 <-
  train(currencycrisis ~ ., 
        data=train4,  
        method = "knn", 
        metric = "ROC", 
        tuneGrid = knn_tune,
        trControl = control_conditions
)
mod_knn2
```


```{r}
# Draw out the best model (the one with the best performance)
mod_knn2$finalModel
```
```{r}
plot(mod_knn2)
```

#2. CART
```{r}
mod_cart <- 
  train(currencycrisis ~ .,
        data = train4,
        method = "rpart",
        metric = "ROC",
        trControl = control_conditions
  )
mod_cart
```

```{r}
plot(mod_cart)
```
```{r}
install.packages("rattle") 
```

```{r}
library(rattle)
```

```{r}
fancyRpartPlot(mod_cart$finalModel)
```

```{r}
print(mod_cart$finalModel)
```


```{r}
tune_cart2 <- expand.grid(cp = c(0.0010281))
mod_cart2 <- 
  train(currencycrisis ~ .,
        data = train4,
        method = "rpart",
        metric = "ROC",
        tuneGrid = tune_cart2,
        trControl = control_conditions
  )
mod_cart2
```

```{r}
fancyRpartPlot(mod_cart2$finalModel)
```

#3. Random Forest

```{r}
install.packages('e1071')
```

```{r}
library(e1071)
```


```{r}
mod_rf <-
  train(currencycrisis ~ .,
        data=train4, # Training data
        method = "ranger", # random forest (ranger is much faster than rf)
        metric = "ROC", # area under the curve
        importance = 'impurity', # For variable importance metric (see below)
        trControl = control_conditions
)
```

```{r}
mod_rf
```

```{r}
rf_tune <- expand.grid(mtry = c(1,3,5,6), splitrule=c("gini", "extratrees"),min.node.size=1)
```


```{r}
mod2_rf <-
  train(currencycrisis ~ .,
        data=train4, # Training data
        method = "ranger", # random forest (ranger is much faster than rf)
        metric = "ROC", # area under the curve
        importance = 'impurity', # For variable importance metric (see below)
        tuneGrid = rf_tune,
        trControl = control_conditions
)
```

```{r}
mod2_rf
```

## Task 6. Compare the output of all three models
```{r}
mod_list <-
  list(
    knn1 = mod_knn,
    knn2 = mod_knn2,
    cart1 = mod_cart,
    cart2 = mod_cart2,
    rf = mod_rf, 
    rf2 = mod2_rf
  )

# Resamples allows us to compare model output
resamples(mod_list)
```

## To Tingjie: print the following graph
```{r}
ml <- dotplot(resamples(mod_list))
ml
```



## Task 7. Test the out-of-sample predictive performance for the best performing model, which is mod2_rf. 

```{r}
View(test4)
```

## Tingjie: use a table to visualize the predictive capability
```{r}
pred <- predict(mod2_rf,newdata = test4)
confusionMatrix(table(pred,test4$currencycrisis))
```



## Task 8. For the Random Forest model, three most important variables are external debt to gni, lagged gdp per capita, and lagged total reserves.


## To Tingjie: print the following graph
```{r}
rf2var <- plot(varImp(mod2_rf),main="Importance of control variables")
rf2var
```





