---
title: "External Debt and Currency Crisis in Emerging Markets"
author: "New Group Project"
date: "April 29, 2019"
header-includes:
    \usepackage{setspace}\doublespacing
output: pdf_document
number_sections: true
fontsize: 12pt
documentclass: article
---


\textbf{Members:}
\begin{itemize}
  \itemsep-.6em
  \item Liumin Chen (NetID: lc1077) - email: lc1077@georgetown.edu\
  \item Tingjie  Meng (NetID: tm1305) - email: tm1305@georgetown.edu\
\end{itemize}
\textbf{Github Repository:} https://github.com/TingjieMeng/New-Group-Project 

# Problem Statement

Debts and credits are crucial both for the expansion of businesses and for economic development. However, when the growth of debts outpaces the growth of income, debts become a time bomb triggering crises. There are two types of debts: domestic debts denominated in the domestic currency and external debts denominated in foreign currencies. The history of financial crises tells us that countries with a high level of external debt are especially vulnerable to capital outflows and thus exchange rate fluctuations. As Ray Dalio describes in <Principles for Navigating Big Debt Crisis>, such an event is called an inflationary depression.

Classic examples of a deflationary depression are the Tukish currency crises in the early 2000s and 2018. Since the 1980s, Turkey has been heavily dependent on foreign investments to fund its huge current account deficits, especially in the capital-intensive construction sector. Thus its economy is fragile to any internal or external shock that can lead to foreign capital outflow on a large scale. In 2001, foreign investors quickly withdrew capital from Turkey amid political tensions and tightening fiscal policies by the Turkish government, which caused the value of Turkish lira to plunge. The history repeats itself. In 2018, due to domestic instability and tightening monetary policies from major central banks, Turkey saw a year-over-year decline of foreign investments by 9,500 basis points. Without sufficient foreign capital to support its weak economy, the Turkish lira soon began drastically losing value, leaving many companies unable to serve their foreign currency debts.

Our project focuses on studying the link between external debt and currency crisis in Emerging Market Economies (EME). As shown in Turkey's crises, over-dependence on external financing can be extremely dangerous for EMEs in the long run. However, EMEs typically face difficulties in issuing debts in local currency because their local financial markets are underdeveloped. Thus they are more likely to hold a significant portion of external debt in the total debt portfolio compared to developed economies. Specifically, we explore factors that strengthen or weaken the link between external debt and currency crisis in EMEs and use the patterns we find to predict the possibility of a currency crisis happening in future years. Our findings can serve as policy guidance for external debt management for central bank policymakers in emerging markets.


# Research Design
## testable hypothesis
Our main hypothesis that countries who have borrowed excessively in foreign currency are more prone to currency crisis risks. Sufficient foreign reserves, a stable political regime, and a strong economy make countries more resilient against currency depreciation risks. Currency crises are more likely to happen when the U.S. Federal Reserve increases the interst rate because that creates more incentives for investors to bring money from EMEs back into the United States.

## dataset
A set of time series data covering 46 developing countries from 1990 to 2015

## data source
Data sources for this study are as follows: the World Bank Open Data, the International Monetary Fund (IMF) Open Data, and the Polity IV Project conducted by the Center for Systemic Peace. All three data sources are publicly available. The World Bank and IMF APIs can be installed in R, and the Polity IV Project data can be conveniently downloaded as a csv package.  

## variable description
Dependent variable: possibility of a currency crisis happening, defined as a nominal depreciation of 15% within a year
Independent variable of interest: externaldebt/GNI
Control variable: GDP per capita, level of democracy, US 3-month treasury yield, foreign reserves/total external debt


# Methods
## data visualization
Data visualization is used for getting a general understanding of the relationships between independent variables and dependent variables. It is also used as case studies of individual countries that experience a classic pattern of currency crises.

## mixed-effect logit model
The logit model in parameterization form is as follows. It is used for modelling the probability of a currency crisis happening given all variables.

## machine learning techniques 
Machine learning here is used for predicting the classification. We use several models including K nearest neighbours, CART, and Random Forest. After comparison, we choose the one model specification with the most accuracy. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include=FALSE, warning=F,error = F,message = F)
```

```{r, cache=TRUE}
require("WDI")
require("IMFData")
require("skimr")
require("readr")
require("tidyverse")
require("ggplot2")
require("dplyr")
require("ggeffects")
require("lme4")
require("caret")
require("rattle")
require("e1071")
```

```{r}
library('WDI')
```

```{r}
WDIsearch('gdp')
```

```{r}
WDIsearch('external.*debt')
```

```{r}
WDIsearch('exchange.*rate')
```

```{r}
Data_WDI = WDI(indicator=c('NEER', 'NY.GDP.PCAP.KD','FI.RES.TOTL.DT.ZS','DT.DOD.DECT.GN.ZS'), country='all', start=1990, end=2018)
head(Data_WDI)
```

```{r}
names(Data_WDI) <- c('iso2c', 'country', 'year', 'nominal.exchange.rate', 'GDP.per.capita', 'total.reserves.outof.total.external.debt', 'external.Debt/GNI')
```


```{r}
getwd()
```

```{r}
Data_polity <- read.csv('/Users/liuminchen/Documents/Currency Crisis/DATASET/politydata.csv')
```

```{r}
WDI_POLITY <- merge(Data_WDI,Data_polity, by=c("iso2c","year"))
```

```{r}
head(WDI_POLITY)
```

```{r}
library(IMFData)
```


```{r}
IFS.available.codes <- DataStructureMethod('IFS') # Get dimension code of IFS dataset
names(IFS.available.codes) # Available dimension code
IFS.available.codes[[1]] # Possible code in the first dimension
CodeSearch(IFS.available.codes, "CL_INDICATOR_IFS", "Treasury")
```

```{r}
queryfilter <- list(CL_FREA = "A", CL_AREA_IFS = "US", CL_INDICATOR_IFS = "FITB_3M_PA")
```


```{r}
databaseID <- "IFS"
startdate = "1980-01-01"
enddate = "2018-12-31"
checkquery = FALSE
```

```{r}
USTREASURY <- CompactDataMethod(databaseID, queryfilter, startdate, enddate, 
    checkquery, tidy = TRUE)
head(USTREASURY)
```

```{r}
library(tidyverse)
```


```{r}
USTREASURY2 <- USTREASURY[,c(1,2)]
head(USTREASURY2)
```

```{r}
names(USTREASURY2) <- c('year', 'US.3m.treasury.yield')
```

```{r}
WDI_POLITY_USTREASURY <- merge(WDI_POLITY, USTREASURY2, by="year")
head(WDI_POLITY_USTREASURY)
```

```{r}
WDI_POLITY_USTREASURY <- WDI_POLITY_USTREASURY[order( WDI_POLITY_USTREASURY[,2] ),]
WDI_POLITY_USTREASURY2 <- WDI_POLITY_USTREASURY[,-c(8,9)]
head(WDI_POLITY_USTREASURY2)
```

## GENERATE DEPENDENT VARIABLE (DUMMY)
```{r}
WDI_POLITY_USTREASURY2$currency.crisis = ifelse((WDI_POLITY_USTREASURY2$nominal.exchange.rate - lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate))/lag(WDI_POLITY_USTREASURY2$nominal.exchange.rate) <= -0.15, 1, 0)
```

## GENERATE INDEPENDENT LAG VARIABLES
```{r}
WDI_POLITY_USTREASURY2$lag.GDP.per.capita = lag(WDI_POLITY_USTREASURY2$GDP.per.capita)
WDI_POLITY_USTREASURY2$lag.total.reserves = lag(WDI_POLITY_USTREASURY2$total.reserves.outof.total.external.debt)
```


```{r}
mean(WDI_POLITY_USTREASURY2$currency.crisis)
```

```{r}
data_clean <- WDI_POLITY_USTREASURY2 %>% drop_na(currency.crisis)
```

```{r}
mean(data_clean$currency.crisis)
```

```{r}
isNA <- apply(is.na(data_clean), 1, any)
data_clean <- data_clean[!isNA, ]
```

```{r}
mean(data_clean$currency.crisis)
```


```{r}
names(data_clean) <- c('year', 'isoc2', 'countryx', 'nominalexchangerate', 'gdppercapita', 'totalreservesoutoftotalexternald', 'externaldebtgni', 'polity2', 'us3mtreasuryyield', 'currencycrisis', 'laggdppercapita', 'lagtotalreserves')
```


```{r}
write.csv(data_clean,'data_clean_for_next_steps.csv')
```


```{r}
df <- read.csv('data_clean_for_next_steps.csv')
```

```{r}
head(df)
```


summary of statistics:
describe the variables
```{r}
skimr::skim(df)
```


# Results
## data visualization
creat df1 to select only China, Argentina, Brazil, Romania, Russia Federation, Sierra Leone, Thailand, Turkey and South Africa. df1 is used to create graph1. display the external debt/gni by 11 countries 

```{r}
df1 <- df%>%
  filter(countryx=="China"|countryx=="Brazil"|countryx=="Romania"|countryx=="Russian Federation"|countryx=="Sierra Leone"|countryx=="Thailand"|countryx=="Turkey"|countryx=="South Africa"|countryx=="Nigeria"|countryx=="Zambia"|countryx=='Ukraine')
```

```{r}
df1$currencycrisis <- factor(df1$currencycrisis)
```

```{r, include=TRUE}
ggplot(df1,aes(year,externaldebtgni))+
  geom_line()+
  ylab("external debt/gni")+
  facet_wrap(~countryx)
```

Is it the higher the external debt, the more often currency crisis happened?
-yes, for Nigeria,Zambia,Thailand,Sierra Leoone.
```{r,fig.height=8,fig.width=15, INCLUDE = TRUE}
p <- ggplot(df1,aes(year,externaldebtgni))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("external debt/gni")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)
  
p
                
```

multicollinearity:
see the relationship between control variable: polity score and gdp per capita
display polity scores by countries:
```{r, INCLUDE = TRUE}
q1 <- ggplot(df1,aes(year,polity2))+
  geom_line()+
  ylab("polity score")+
  facet_wrap(~countryx)
q1
```
display GDP per capita by countries
```{r, INCLUDE = TRUE}
q2 <- ggplot(df1,aes(year,laggdppercapita,color="red"))+
  geom_line()+
  facet_wrap(~countryx)
q2
```

multicolinear?
No. multicolinearity doesn't seem to exist.
it make sense: Nigeria, Romania -> The more "democratic", the higher lagged gdp per capita
not make sense: China, Brazil,Russia, Ukraine, Sierra Leone, South Africa, Thailand, Turkey
1) The polity score might not a strong factor reflecting political stability
2) The polity score might not be strongly asscoiated with some countries' currency crisis.
```{r,fig.height=8,fig.width=15, INCLUDE = TRUE}
ggplot(df1,aes(x=year))+
  geom_line(aes(year,laggdppercapita),color="steel blue")+
  geom_line(aes(year,(polity2+10)*500),color="dark red")+
  scale_y_continuous(name="lagged gdp per capita", sec.axis=sec_axis(~./500-10, name="polity score"))+
  theme(axis.title.y.left=element_text(color="steel blue"),
    axis.text.y.left=element_text(color="steel blue"),
    axis.title.y.right=element_text(color="dark red"),
    axis.text.y.right=element_text(color="dark red"))+
  facet_wrap(~countryx)
```

external shock: How U.S. interest rate influence EMEs' currency crisis?

```{r}
 df2 <- df%>%
  filter(currencycrisis==1)
```
Count the number of currency crisis of all selected countries by years
```{r}
ggplot(df2,aes(x=year))+
geom_bar(mapping = aes(x = year, y = currencycrisis), stat = "identity") 
```
Display the U.S. interest rate by years
```{r}
ggplot(df2,aes(x=year))+
  geom_line(aes(year,us3mtreasuryyield))+
  ylab("US Interest Rate")
```
The trends of US interest rate and # of countries undergoing currency crisis appear to move in the same direction approximately. 
US interest rate as an external shock has an impact on EMEs' currency crisis.

```{r, INCLUDE = TRUE}
ggplot(df2,aes(x=year))+
  geom_line(aes(year,us3mtreasuryyield*2.5),color="dark red")+
  geom_bar(mapping = aes(x = year, y = currencycrisis), stat = "identity",fill="steel blue",alpha=0.4)+
  scale_y_continuous(name="# of countries with currency crisis", sec.axis=sec_axis(~./2.5, name="US interest rate"))+
  theme(axis.title.y.left=element_text(color="steel blue"),
    axis.text.y.left=element_text(color="steel blue"),
    axis.title.y.right=element_text(color="dark red"),
    axis.text.y.right=element_text(color="dark red"))
```

polity score and currency crisis by country: no direct relationship found
```{r,fig.height=8,fig.width=15, INCLUDE = TRUE}
 ggplot(df1,aes(year,polity2))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("polity score")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)

```

The lower the foreign reserves, the more currency crisis happened.
strong relationship!
```{r,fig.height=8,fig.width=15, INCLUDE = TRUE}
 ggplot(df1,aes(year,totalreservesoutoftotalexternald))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("total foreign reserves out of total external debt")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)
```


## LOGIT REGRESSION

## mixed-effect logit model
The logit model shows that there is a positive correlation between our independent variable of interest and the dependent variable, and such correlation is statistically significant at the 0.1% level. This supports our main hypothesis that countries who have borrowed excessively in foreign currency are more prone to currency crisis risks. 
In terms of control variables, the model shows a negative correlation between the ratio of foreign reserves to a country’s total external debt in the previous year and the dependent variable. There is also a negative correlation between GDP per capita in the previous year and the probability of a currency crisis happening. However, the correlations for both these two control variables are not statistically significant. There is a positive correlation between US 3-month treasury yield and the dependent variable, but again this correlation is not significant. Generally, our finding is consistent with our theory, that countries with stronger economies (higher GDP per capita) and sufficient foreign reserves are more resilient against currency crises. Besides, in terms of external factors, currency risks are particularly high during times when the US Federal Reserve sets the interest rate high.  
In our hypothesis, countries which have higher polity scores typically have more transparent and efficient public management systems. That enables those countries to deal with exchange rate problems more effectively. However, the positive coefficient for polity score shown here is in contrast with our expectation. This is most likely due to the dataset issue, that the polity score is not the best indicator for level of democracy. Countries like China and Brazil have low polity scores but have stable exchange rates.
There is no good equivalent to R-squared for logistic regression models. One alternative is to look at the log likelihood, which is -306.2 in this case. Another alternative is to look at the AIC, which is 628.4 here. Both the log likelihood and AIC are relative measures of model fit. A lower value of the AIC or higher value of log likelihood suggests a relatively "better" model.

```{r}
final_data <- read.csv('data_clean_for_next_steps.csv')
```

```{r}
library(lme4)
m <- glmer(currencycrisis ~ laggdppercapita + lagtotalreserves + externaldebtgni + polity2 + us3mtreasuryyield + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m, corr = FALSE)
```

```{r}
summary(m)
```

```{r}
plot(fitted(m), residuals(m), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
```



```{r}
library(ggeffects)
margin_eff <- ggpredict(m, term = "externaldebtgni")
margin_eff
```


```{r}
plot(margin_eff)
```

```{r}
margin_eff2 <- ggpredict(m, term = "lagtotalreserves")
margin_eff2
```

```{r}
plot(margin_eff2)
```

```{r}
margin_eff3 <- ggpredict(m, term = "us3mtreasuryyield")
margin_eff3
```

```{r}
plot(margin_eff3)
```

```{r}
margin_eff4 <- ggpredict(m, term = "laggdppercapita")
margin_eff4
```

```{r}
plot(margin_eff4)
```


```{r}
margin_eff5 <- ggpredict(m, term = "polity2")
margin_eff5
```

```{r}
plot(margin_eff5)
```


```{r}
library(ggplot2)
```


```{r}
m3 <- glmer(currencycrisis ~  externaldebtgni + polity2 + us3mtreasuryyield + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m3, corr = FALSE)
summary(m3)
```

```{r}
m4 <- glmer(currencycrisis ~  lagtotalreserves + externaldebtgni + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m4, corr = FALSE)
summary(m4)
```

```{r}
m5 <- glmer(currencycrisis ~  externaldebtgni + (1|year) + (1|countryx), data = final_data, family = 'binomial')
print(m5, corr = FALSE)
summary(m5)
```

```{r}
library(ggpubr)
```

Figure 6 shows the partial dependency ...
```{r,include= TRUE, fig.align="center", fig.width=14,fig.height=8}
#partial dependency plot 
partial <- ggarrange(margin_eff, margin_eff2, margin_eff3, margin_eff4, margin_eff5)
partialmerge <- annotate_figure(partial,
                top = text_grob("Partial dependency plot \n figure 6", color = "dark grey", face = "bold", size = 14))+
  ggsave("partialmerge.jpg")
partialmerge
```

## MACHINE LEARNING 

We split the data into a training and test dataset, partitioning data before 2014 into the training data, and holding out data in and after 2014 as a test set. normalize the scale. KNN, CART, RF, Compare the output of all three models. ## Test the out-of-sample predictive performance for the best performing model, which is mod2_rf. use a table to visualize the predictive capability.  For the Random Forest model, three most important variables are external debt to gni, lagged gdp per capita, and lagged total reserves.


```{r}
getwd()
```

```{r}
ml_data <- read.csv(file = "/Users/liuminchen/Documents/Currency Crisis/data_clean_for_next_steps.csv")
```



```{r}
test = ml_data %>% filter(year >= 2014)
train = ml_data %>% filter(year < 2014)
```



```{r}
library(recipes)
```

```{r}
train2 <- 
  train %>% mutate_if(is.factor, as.numeric)
```

```{r}
test2 <- 
  test %>% mutate_if(is.factor, as.numeric)
```

```{r}
rcp <-
  recipe(countryx~.,train2) %>%
  step_range(countryx, polity2,us3mtreasuryyield,laggdppercapita, lagtotalreserves, externaldebtgni) %>%  # Normalize scale
  prep()
# Apply the recipe to the training and test data
train3 <- bake(rcp,train2)
test3 <- bake(rcp,test2)
```

```{r}
train4 <- train3[,c(4,8,9,10,11,12,13)]
test4 <- test3[,c(4,8,9,10,11,12,13)]
```


```{r}
library(caret)
library(rattle)
library(e1071)
set.seed(1988) # set a seed for replication purposes 
folds <- createFolds(train4$currencycrisis, k = 5) # Partition the data into 10 equal folds
sapply(folds,length)
```

```{r}
control_conditions <- 
  trainControl(method='cv', # K-fold cross validation
               summaryFunction = twoClassSummary, # Need this b/c it's a classification problem
               classProbs = TRUE, # Need this b/c it's a classification problem
               index = folds # The indices for our folds (so they are always the same)
  )
```





```{r}
train4$currencycrisis <- factor(train4$currencycrisis)
test4$currencycrisis <- factor(test4$currencycrisis)
```

```{r}
levels(train4$currencycrisis) <- c("no", "yes")
levels(test4$currencycrisis) <- c("no", "yes")
```

```{r}
mod_knn <-
  train(currencycrisis ~ ., # Equation (outcome and everything else)
        data=train4, # Training data 
        method = "knn", # K-Nearest Neighbors Algorithm
        metric = "ROC", # area under the curve
        trControl = control_conditions
  )
mod_knn
```
```{r}
knn_tune = expand.grid(k = c(1,3,5,6,7,10))
knn_tune
```

```{r}
mod_knn2 <-
  train(currencycrisis ~ ., 
        data=train4,  
        method = "knn", 
        metric = "ROC", 
        tuneGrid = knn_tune,
        trControl = control_conditions
)
mod_knn2
```


```{r}
# Draw out the best model (the one with the best performance)
mod_knn2$finalModel
```
```{r}
plot(mod_knn2)
```

```{r}
mod_cart <- 
  train(currencycrisis ~ .,
        data = train4,
        method = "rpart",
        metric = "ROC",
        trControl = control_conditions
  )
mod_cart
```

```{r}
plot(mod_cart)
```


```{r}
fancyRpartPlot(mod_cart$finalModel)
```

```{r}
print(mod_cart$finalModel)
```


```{r}
tune_cart2 <- expand.grid(cp = c(0.0010281))
mod_cart2 <- 
  train(currencycrisis ~ .,
        data = train4,
        method = "rpart",
        metric = "ROC",
        tuneGrid = tune_cart2,
        trControl = control_conditions
  )
mod_cart2
```

```{r}
fancyRpartPlot(mod_cart2$finalModel)
```


```{r}
mod_rf <-
  train(currencycrisis ~ .,
        data=train4, # Training data
        method = "ranger", # random forest (ranger is much faster than rf)
        metric = "ROC", # area under the curve
        importance = 'impurity', # For variable importance metric (see below)
        trControl = control_conditions
)
```

```{r}
mod_rf
```

```{r}
rf_tune <- expand.grid(mtry = c(1,3,5,6), splitrule=c("gini", "extratrees"),min.node.size=1)
```


```{r}
mod2_rf <-
  train(currencycrisis ~ .,
        data=train4, # Training data
        method = "ranger", # random forest (ranger is much faster than rf)
        metric = "ROC", # area under the curve
        importance = 'impurity', # For variable importance metric (see below)
        tuneGrid = rf_tune,
        trControl = control_conditions
)
```

```{r}
mod2_rf
```

```{r}
mod_list <-
  list(
    knn1 = mod_knn,
    knn2 = mod_knn2,
    cart1 = mod_cart,
    cart2 = mod_cart2,
    rf = mod_rf, 
    rf2 = mod2_rf
  )

# Resamples allows us to compare model output
resamples(mod_list)
```

```{r, INCLUDE = TRUE}
dotplot(resamples(mod_list))
```


```{r, INCLUDE = TRUE}
pred <- predict(mod2_rf,newdata = test4)
confusionMatrix(table(pred,test4$currencycrisis))
```


```{r, INCLUDE = TRUE}
plot(varImp(mod2_rf))
```


# Conclusion
## Lessons Learned
The first and most straightforward policy implication is that countries should be cautious about their level of external debt, especially emerging market economies (EME). Secondly, EMEs can take multiple measures to increase thier resilience against currency crises.Those include stocking foreign reserves, maintaining the credibility of central banks in the international capital market, and developing a strong economy. The third policy implication is that currency crisis risks are particularly high in times of Fed rate hikes. This reminds policymakers from national central banks and international organizations like the IMF that EMEs’ external debts need to be more closely monitored in times of interest rate volatility. 

## Next steps
First, as mentioned in the results analysis part, polity score published by Center for Systemic Peace measures the level of democracy, not the political regime stability. That introduces bias into our model. Besides, potentially there can be more variables to be included in the model, like the economic structure and exchange rate regime. Third, our dataset only includes data until 2015, which means that currency crises in Turkey and Argentina in 2018 are not in our sample. In future research, it will be meaningful to see if our findings remain the same when a larger sample is used.




