---
title: "Analyais on external debt and currency crisis in Emerging markets"
author: "Liumin Chen & Tingjie Meng"
date: "April.29 2019"
output:
  beamer_presentation: default
  pdf_document: default
---

```{r}
knitr::opts_chunk$set(echo=FALSE)
```

## Problem statement and Background

- EMEs (emerging markets) are vulnerable: inflationary depression
1) Countries with a high level of external debt are especially vulnerable to capital outflows and thus exchange rate fluctuations
- the link between external debt and currency crisis in Emerging Market Economies (EME)
1) EMEs are more likely to hold a significant portion of external debt compared to developed economies. 
2) What factors strengthen or weaken the link between external debt and currency crisis in EMEs?
3) How to predict the possibility of a currency crisis happening in future years?
4) Policy guidance for external debt management for central bank policymakers in emerging markets

## Approach (page 1)

- Data description
1) the World Bank Open Data
2) the International Monetary Fund (IMF) Open Data
3) the Polity IV Project conducted by the Center for Systemic Peace

- Dependent variables (DV): Currency crisis happened in a country by years

- Control variables: External debt/GNI, U.S. interest rate, foreign exchange reserves, polity score, GDP per capita

- Data access: Public data 

- Data limitation
1) More difficult to obtain macroeconomic data on less developed countries because of bureaucratic mismanagement and officials' incentive to inflate their economic data 
2) Polity score is not a perfect proxy for political stability

## Methods 

- Data visualizations - ggplot2
1) explore the connection between DV and control variables
2) explore the connections among control variables

- Logit model  

- Machine learning 

## Result: data visualizations  (page 1)
- connection between external debt and currency crisis
(graph 1)
- connection between polity score and gdp per capita
(graph 2)
- How U.S. interest rate influence EMEs' currency crisis?
(graph 3)

## Result: data visualization (page 2)
- polity score and currency crisis by country: not significant
(graph 4)
- The lower the foreign reserves, the more currency crisis happened
(graph 5)

## Result: Logit model (page 3)

## Result: Machine learning (page 4)

## Lessons learned

```{r, echo=FALSE}
require(tidyverse)
require(ggplot2)
require(dplyr)
```



```{r, echo=FALSE}
df <- read.csv("data_clean_for_next_steps.csv")
df1 <- df%>%
  filter(countryx=="China"|countryx=="Brazil"|countryx=="Romania"|countryx=="Russian Federation"|countryx=="Sierra Leone"|countryx=="Thailand"|countryx=="Turkey"|countryx=="South Africa"|countryx=="Nigeria"|countryx=="Zambia"|countryx=='Ukraine')
df1$currencycrisis <- factor(df1$currencycrisis)
ggplot(df1,aes(year,externaldebtgni))+
  geom_line()+
  geom_point(aes(year,color=currencycrisis))+
  ylab("external debt/gni")+
  scale_color_manual(values=c("black","red"))+
  facet_wrap(~countryx)+
  ggsave("p.jpg")
```

```{r, cars, fig.cap="A scatterplot.", echo=FALSE}
plot(cars)
```

